Algebraic Groups\vspace{-1ex}

### 2019-11-06 Algebraic Geometry/Algebraic-Geometry.tex
update

\subsection{Homogenization/dehomogenization}
Given a homogenous polynomial $f(x_0,\dots,x_n)$ of degree $d$, we can \textit{dehomogenize} things by setting $x_0=1$. For instance $x_0^2+x_1^2+x_2^2\mapsto 1+x_1^2+x_2^2$.
Given a polynomial $g(x_1,\dots,x_n)$, we just look at 
\[\tilde g(x_0,\dots,x_n)=x^{\deg g}g(x_1/x_0,\dots,x_n/x_0).\]
One can see relatively easily that $\tilde g$ is a homogeneous polynomial of degree $d$.

@@

\end{document}

>>>>

\subsection{Homogenization/dehomogenization}
Given a homogenous polynomial $f(x_0,\dots,x_n)$ of degree $d$, we can \textit{dehomogenize} things by setting $x_0=1$. For instance $x_0^2+x_1^2+x_2^2\mapsto 1+x_1^2+x_2^2$.
Given a polynomial $g(x_1,\dots,x_n)$, we just look at 
\[\tilde g(x_0,\dots,x_n)=x_0^{\deg g}g(x_1/x_0,\dots,x_n/x_0).\]
One can see relatively easily that $\tilde g$ is a homogeneous polynomial of degree $d$.

@@

\section{November 6, 2019}
Today we are going to focus on the pickles and not the meat. Go Max.

\subsection{Questions and Thoughts}
First things first, we wanted to talk about the problem above. It is easy to see that homogenizing and then dehomongenizing gives you the same polynomial.
For the other direction, it seems that a (homogeneous) polynomial comes back to itself if it is not divisible by a power of $x_0$.

What is happening geometrically? Dehomogenizing is intersecting with an open affine subset. Homogenizing is taking the closure.
\begin{rmk}
	The idea came up that taking the closure might be the same as taking the cone containing the original thing. But consider 
	the hyperbola $xy-1$ and its homogenization $xy-z^2$. Now $Z=Z(xy-z^2)$ is closed and contains $H=Z(xy-1)$, so it contains $\bar H$.

	Conjecture: $Z\cong \P1\subseteq\P2$ as a conic. How can we realize this as a map from $\P1$ to $\P2$? Well,
	we need a map 
	\[(x_0,x_1)\mapsto(f_0(x_i),f_1(x_i),f_2(x_i))\]
	where the $f_i$ are homogeneous of the same degree and have no common zeros in $\P1$.
\end{rmk}
\begin{rmk}
	More generally, given varieties $X$ and $Y$, a \textbf{rational map} from $X$ to $Y$ $f:X\dashrightarrow Y$ is an equivalence class of morphisms $U\to Y$ where $U$ is open in $X$ 
	and $U\xrightarrow{\alpha} Y$ and $V\xrightarrow{\beta}X$ are equvialent if $\alpha|_{U\cap V}=\beta|_{U\cap V}$.

	Then we can choose $f_0,\dots,f_m\in k[x_0,\dots,x_n]$ that are homogeneous of degree $d$ and these coordinate functions give us a rational map (in fact, all the rational maps) from $\P n$ to $\P m$.
\end{rmk}
\begin{prob}
	Find a morphism $f:\P1\to\P2$ whose image is $Z(xy-z^2)$.
\end{prob}
I am thinking $x_0^2$, $x_1^2$, and $x_0x_1$.
\begin{prob}
	Suppsoe that $Q$ is a non-singular quadratic form in three variables, so $Z(Q)\subseteq \P2$. Then show that $Z(Q)\cong \P1$.
	\textit{Hint: Linear changes of coordinates are algebraic!}
\end{prob}
\end{document}

### 2019-11-06 Algebraic Groups/Algebraic-Groups.tex
update

>>>>

\section{November 6th, 2019}
Today we are going to prove some more things about representation theory. Down the pipe somewhere we will hope to talk about Tanakka duality, 
which gives us that Morita equivalence (is this still what it's called?) of two algebraic groups gives us the groups are isomorphic.

Given an algebraic group $G/k$, we could consider representations, which were either a group scheme morphism $G\to \GL(V)$ or else a coaction $V\xrightarrow{\sigma} \Gamma(G)\otimes_k V$.
Then we said that $G$ was linear reductive if and only if every representation is completely reducible,

\begin{ex}
	If $G=\Gm$, then every irreducible representation is one-dimensional: $\Gm\to\Gm$ sends $t\mapsto t^\alpha$.
\end{ex}
A natural question one may ask: what can we say in general?
\subsection{The regular representation}
The regular representation is $\Gamma(G)$ with coaction given by comultiplicaton.
\begin{prop}
	And finite dimentional representation $V$ of $G$ embeds $V\subseteq \Gamma(G)^{\oplus n}$.
\end{prop}
\begin{prf}
	The coaction gives us a map 
	\[V\xrightarrow{\sigma}\Gamma(G)\otimes V=\Gamma(G)\otimes_k V_{\text{vs}}\cong \Gamma(G)^{\oplus\dim V}\]
	where $V_{vs}$ is the underlying vector space of $V$. Then the claim is that the overall map is injective and that it is a map of $G$-reps.

	The latter property is almost tautological by the property of a coaction. Injectivity follows from the fact that 
	\[(e^\ast\otimes\id)\circ\sigma\]
	is an isomorphism $V\to k\otimes V$ (again by one of the coaction axioms), so $\sigma$ is injective.
\end{prf}
\begin{prop}
	If $V$ is a finite dimensional faithful representation, then every other finite-dimensional representation can be obtained from $V$ by direct sums, tensors, duals, subrepresentations, and quotients.
\end{prop}
\begin{rmk}
	More precisely, $W\subseteq (V\oplus V^\vee)^{\otimes n}$.
\end{rmk}
\begin{prf}
	$G\subseteq\GL(V)$ is a closed subgroup. Notice that since we have an injective map into $\GL(V)$, $G$ is already a linear algebraic group.
\end{prf}
\begin{thm}[Peter-Weyl]
	If $G$ is linear reductive, then 
	\[\Gamma(G)=\oplus_{\text{irr $V$}}(V\otimes V^\vee)\]
\end{thm}
\begin{rmk}
	Notice that this is as two-sided representations, but if we only want to look at (say) left representations, we give $V^\vee$ the trivial 
	representation structure and we get the more familiar 
	\[\Gamma(G)=\oplus_{\text{irr $V$}} V^{\oplus\dim V}\]
	and then the formula from finite groups:
	\[|G|=\sum_{\text{irr $V$}} (\dim V)^2\]
\end{rmk}
As an example, consider $\Gm=\Spec k[t]_t$ and $k[t]_t=\oplus_{n\in\bbZ}k\langle t^n\rangle$. As an exercise, thing about what happens for $\SL_2$.

\subsection{Stabilizers of subspaces}
If $G$ is an algebraic group over $k$ and $V$ is a $G$ representation, let $W\subseteq V$ be a subspace. We want a subgroup $G_W$ which plays a similar role as the stabilizer in group theory.

Consider $G$ as a functor from $\Alg_k\subseteq \Sch/k\to\Grp$ taking $T\mapsto\Hom(T,G)$. Then for a $k$-algebra $R,$ $G(R)$
acts on $V_R=V\otimes R$.
\begin{prop}
	The functor $\Alg_k\to\Grp$ sending 
	\[R\mapsto \{g\in G(R)|gW_R=W_R\}\]
	is representable by a subgroup of $G$, which we will denote $G_W$.
\end{prop}
\begin{rmk}
	The idea here is to fix a basis $e_i$ of $W$ and complete them to a basis of $V$ by appending $f_i$. Then take the coaction $V\to\Gamma(G)\otimes V$ and consider the image 
	\[f_k\mapsto \sum_{i\in I}a_{ki}\otimes e_i+\sum_{j\in J}a_{kj}\otimes f_j\]
	suppose we ahve $g\in G(R)=\Hom(\Spec R, G)$. This gives us a map in $\Hom(\Gamma(G),R)$. Then using the action on $V\otimes R$:
	\[g\cdot(f_k\otimes 1)=\sum g(a_{ki})\otimes e_i+\sum g(a_{kj})\otimes f_j\in V_R\]
	and we see that this is actually in $W_R$ exactly when $g(a_{kj})=0$ for all $j$.

	But then this means that $g$ lands in $V(a_{kj})$, the vanishing of the functions $a_{kj}\otimes 1\in\Gamma(G)_R$, so we get that 
	the closed subscheme $V(a_{kj})$ represents the functor $G_W$.
\end{rmk}

For the converse, let $G$ be an algebraic group over $k$ and let $H\subseteq G$ be a subgroup. Then there exists a finite dimensional representation $V$ of $G$ and 
a $L\subseteq V$ a one-dimensional subspace such that $H=G_L$.

As an example, look at $G=\SL_2$ acting on $k^2$. Let $L=\langle 1,0\rangle$. We can see easily that $L$ is preserved by $g\in G$ if and only if the lower-left coordinate of $g$ is zero.
Now if we take $\Gm=\operatorname{diag}(t,t^{-1})\subseteq\SL_2$, consider the representation $\langle xy\rangle\subseteq \operatorname{Sym}^2k^2$. This isn't quite it but maybe you can work it out!
